---
title: "assignment04"
format: 
  html: 
    self-contained: true
editor: source
author: "Katie Ward and Ben Burnley"
---

## Set-Up

```{r}
#| echo: false
#| message: false
 
library (tidyverse)
library (haven)

# this file is in the .ignore
data <- read_sav("data/ATP_W101.sav")

```

Our data are from [Pew Research Center's American Trends Panel Wave 101](https://www.pewresearch.org/american-trends-panel-datasets/). This survey asks a battery of traditional political questions and this particular wave focused on political priorities. The survey was in the field January 10 -- January 17, 2022.

\newline

## Chart 1

```{r}
#| code-fold: true
#| code-summary: "Show data cleaning code"
#| message: false

# create data for first chart
chart1data <- data %>%
  filter(F_PARTYSUM_FINAL == 1 | F_PARTYSUM_FINAL == 2 ) %>%
  mutate(parties = case_when(F_PARTYSUM_FINAL == 1 ~"Republican/Lean Repub.", 
                             F_PARTYSUM_FINAL == 2 ~ "Democrat/Lean Dem.", 
                             TRUE~NA_character_)) %>%
  mutate(religiousactivity = case_when(F_ATTEND ==1 ~ "More than once a week", 
                                       F_ATTEND == 2 ~ "Once a week", 
                                       F_ATTEND == 3 ~ "Once or twice a month", 
                                       F_ATTEND ==4 ~ "A few times a year", 
                                       F_ATTEND ==5 ~ "Seldom", 
                                       F_ATTEND ==6 ~ "Never", 
                                       TRUE~NA_character_ ))

# create levels for religious activity variable
rel_levels = c("Never", "Seldom", "A few times a year", "Once or twice a month", "Once a week", "More than once a week")  

# factor the religious activity variable
chart1data <- chart1data %>% 
  mutate(religiousactivity = factor(religiousactivity, levels = rel_levels))
```

```{r}
#| code-fold: true
#| code-summary: "Show code for chart"
#| message: false

chart1 <- chart1data %>% 
  filter(!is.na(religiousactivity)) %>%
  ggplot(aes(religiousactivity, fill=parties)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = c("darkblue", "darkred")) +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Religious Service Attendance by Party Identification",
    subtitle = "Aside from weddings and funerals, how often do you attend religious services?",
    x = "Attendance Level",
    y = "Number of Respondents",
    caption = "Source: Pew Data Center American Trends Panel Wave 101",
    fill = "Party ID"
  ) +
  theme(
    axis.text.x = element_text(angle = 70), 
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(face = "italic")
  )

chart1
```

The graph above shows the relationship between party identification and religious service attendance. A large number of the respondents who identify as Democrats report attending religious services "never" or "seldom". For respondents that identify as Republican, there is a greater number that attend more frequently. The difference between the parties is not as large as expected in those who attend religious services once a week to once a month.

\newline

## Chart 2

```{r}
#| code-fold: true
#| code-summary: "Show data cleaning code"
#| message: false

chart2data <- chart1data %>%
  filter(SATIS_W101 != 99) %>%
  mutate(satisfied = ifelse(SATIS_W101==1, 1, 0)) %>%
  mutate(notsatisfied = ifelse(SATIS_W101==2, 1, 0)) %>%
  group_by(parties) %>%
  summarize(satisfied = sum(satisfied), notsatisfied = sum(notsatisfied))

```

```{r}
#| code-fold: true
#| code-summary: "Show code for chart"
#| message: false

chart2 <- chart2data %>%
  ggplot(aes (x = "", y = satisfied, fill=parties)) +
  geom_col() +
  coord_polar("y", start=0) +
  scale_fill_manual(values = c("darkblue", "darkred")) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(), 
    axis.title = element_blank(),
    axis.line = element_blank(),
    plot.subtitle = element_text(size = 8),
    plot.title = element_text(face = "bold"),
  ) + 
  labs(
    title = "Satisfication Rate by Party",
    subtitle = "Respondents who said they were satisfied with the way things are going in this country.",
    caption = "Source: Pew Data Center American Trends Panel Wave 101",
    fill = "Party ID")

chart2
  
```

The pie chart above gives the breakdown by party of respondents that say they are satisfied with the way things are going in the country. 79% of respondents who are satisfied with the country identify as Democrats compared to only 21% who identify as Republicans. Given that Democrats control both chambers of Congress and the Presidency during the time of this survey, this is not necessarily surprising. The proportions are likely not representative of the population as a whole given this is an unweighted proportion of respondents in our sample.

\newline

## Chart 3

```{r}
#| code-fold: true
#| code-summary: "Show data cleaning code"
#| message: false

chart3data <- data %>%
  filter(F_AGECAT != 99) %>%
  mutate(age = case_when(F_AGECAT ==1 ~"18-29", 
                         F_AGECAT==2 ~ "30-49", 
                         F_AGECAT==3 ~ "50-64", 
                         F_AGECAT ==4 ~ "65+", 
                         TRUE~NA_character_ )) %>%
  group_by(age) %>%
  summarize(ss = sum(PRIPRES_SC_W101==1, na.rm = TRUE),
            climate = sum(PRIPRES_GCC_W101==1, na.rm = TRUE), 
            crime = sum(PRIPRES_CRI_W101==1, na.rm = TRUE), 
            immigration = sum(PRIPRES_IMM_W101==1, na.rm = TRUE)) %>%
  pivot_longer(cols = 2:5, names_to = "Policy", values_to = "count")

ages <- data %>% 
  filter(F_AGECAT != 99) %>%
  mutate(age = case_when(F_AGECAT ==1 ~"18-29", 
                         F_AGECAT==2 ~ "30-49", 
                         F_AGECAT==3 ~ "50-64", 
                         F_AGECAT ==4 ~ "65+", 
                         TRUE~NA_character_ )) %>% 
  group_by(age) %>% 
  summarize(total = n())

chart3data2 <- chart3data %>% 
  left_join(ages, by = "age") %>% 
  mutate(percent = count/total)

```

```{r}
#| code-fold: true
#| code-summary: "Show graph code"
#| message: false

chart3 <- chart3data2 %>%
  ggplot(aes(age, percent, color = Policy)) +
  geom_point(size = 3) +
  geom_line(aes(group = Policy)) + 
  theme_minimal() +
  scale_color_discrete(labels = c("Climate Change", "Crime", "Immigration", "Social Security"))+
  labs(
    title = "Top Policy Priorities by Age",
    subtitle = "Percentage of respondents who selected various policy priorities by each age group",
    x = "Age Cohort",
    y = "Percent of Respondents",
    caption = "Source: Pew Data Center American Trends Panel Wave 101"
  ) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 8)
  )

chart3  

```

The chart above shows the percentage of respondents who rank a selection of policy areas as the "top priority". The policy areas we selected -- climate change, crime, immigration, and social security -- represent only a portion of the policy areas respondents could rank. We selected these four because of their lack of overlap. Climate change is the top priority among the youngest cohort with over 27% ranking this as their top priority. This policy topic decreases in relevance in the older cohorts. Crime, immigration, and Social Security all increase in importance as you move up the age cohorts, with Social Security unsurprisingly being of top importance to the oldest cohorts.

## Chart 4

```{r}
#| code-fold: true
#| code-summary: "Show cleaning code"
#| message: false

chart4data <- data %>%
  filter(F_AGECAT != 99) %>%
  mutate(age = case_when(F_AGECAT ==1 ~"18-29", 
                         F_AGECAT==2 ~ "30-49", 
                         F_AGECAT==3 ~ "50-64", 
                         F_AGECAT ==4 ~ "65+", TRUE~NA_character_ )) %>%
  filter(F_PARTYSUM_FINAL == 1 | F_PARTYSUM_FINAL == 2 ) %>%
  mutate(parties = case_when(F_PARTYSUM_FINAL == 1 ~"Republican/Lean Repub.", 
                             F_PARTYSUM_FINAL == 2 ~ "Democrat/Lean Dem.", 
                             TRUE~NA_character_)) 

```

```{r}
#| code-fold: true
#| code-summary: "Show graph code"
#| message: false

chart4 <- chart4data %>%
  ggplot (aes(age, parties, color = parties)) +
  geom_jitter(alpha=0.3) +
  scale_color_manual(values = c("darkblue", "darkred")) +
  theme_minimal()+
  theme (axis.text.y = element_blank(), 
         legend.position = "bottom",
         plot.title = element_text(face = "bold"),
         axis.title.y = element_blank()) +
  labs(
    title = "Party Representation by Age",
    subtitle = "Party identification of survey respondents by age group",
    x = "Age Cohort",
    caption = "Source: Pew Data Center American Trends Panel Wave 101",
    color = "Party ID"
    )

chart4
```

For this final chart, we attempted a novel representation of our sample, with each dot representings one respondent in our sample. We've separated them out by party identification and age cohort. The chart shows several interesting trends. First, there are more Democrats in our sample than Republicans, illustrating concerns of differential non-response in polling. Additionally, though the age cohorts are balanced across party lines, there are very few young people in the survey overall.
